The effect of climate forcing on numerical simulations of the Cordilleran ice sheet at the Last Glacial Maximum

### 2014-03-27 reviews/disc-ac1.tex
Minor changes for latex2rtf compatibility
Add more corrections from Irina

\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex]{color}
\usepackage{doi}

@@

Also, the new Fig.~8 (temperature density maps) now use bilinearly-interpolated temperature maps, rather than the original data (as presented in Fig.~2--4). This results in smoother density plots than those presented in the discussion paper (Fig.~AC4). In a second step, we apply a lapse-rate correction to project all reanalysis data onto the WorldClim (higher resolution) topography (Fig.~AC5). It now becomes clear that much of the discrepancies observed were due to lapse-rate effects.

@@

\begin{figure}
    \center
    \includegraphics[width=70mm]{cordillera-climate-tempdiff+tcdc}
    \caption{Summer temperature difference maps, as in paper Fig.~9.}
\end{figure}

\begin{figure}
    \center
    \includegraphics[width=70mm]{cordillera-climate-tempdiff+lr+tcdc}
    \caption{Summer temperature difference maps, after applying a lapse-rate correction of 6\,{$^\circ$}C\,km$^{-1}$.}
\end{figure}

\begin{figure}
    \center
    \includegraphics[width=70mm]{cordillera-climate-tempdiff+lr}
    \caption{Summer temperature difference maps, after applying a lapse-rate correction of 6\,{$^\circ$}C\,km$^{-1}$ and an updated colour mapping.}
\end{figure}

\begin{figure}
    \center
    \includegraphics[width=70mm]{cordillera-climate-tempheatmap}
    \caption{Summer temperature density maps, after bilinear interpolation.}
\end{figure}

\begin{figure}
    \center
    \includegraphics[width=70mm]{cordillera-climate-tempheatmap+lr}
    \caption{Summer temperature density maps, after bilinear interpolation and lapse-rate correction.}
\end{figure}

\begin{figure}
    \center
    \includegraphics[width=70mm]{cordillera-climate-precheatmap}
    \caption{Winter precipitation density maps, after bilinear interpolation.}
\end{figure}

>>>>

\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex]{color}

@@

Also, the new Fig.~8 (temperature density maps) now uses bilinearly-interpolated temperature maps, rather than the original data (as presented in Fig.~2--4). This results in smoother density plots than those presented in the discussion paper (Fig.~AC4). In a second step, we apply a lapse-rate correction to project all reanalysis data onto the WorldClim (higher resolution) topography (Fig.~AC5). It now becomes clear that much of the discrepancies observed were due to lapse-rate effects.

@@

\begin{figure}
    \centering
    \includegraphics[width=70mm]{cordillera-climate-tempdiff+tcdc}
    \caption{Summer temperature difference maps, as in paper Fig.~9.}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=70mm]{cordillera-climate-tempdiff+lr+tcdc}
    \caption{Summer temperature difference maps, after applying a lapse-rate correction of 6\,{$^\circ$}C\,km$^{-1}$.}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=70mm]{cordillera-climate-tempdiff+lr}
    \caption{Summer temperature difference maps, after applying a lapse-rate correction of 6\,{$^\circ$}C\,km$^{-1}$ and an updated colour mapping.}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=70mm]{cordillera-climate-tempheatmap}
    \caption{Summer temperature density maps, after bilinear interpolation.}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=70mm]{cordillera-climate-tempheatmap+lr}
    \caption{Summer temperature density maps, after bilinear interpolation and lapse-rate correction.}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=70mm]{cordillera-climate-precheatmap}
    \caption{Winter precipitation density maps, after bilinear interpolation.}
\end{figure}

### 2014-03-27 reviews/disc-ac2.tex
Add more corrections from Irina

where $\nu=1\times10^{21}$\,Pa\,s is mantle viscosity, $\rho_r = 3300$\,kg\,m$^{-3}$ is lithosphere density, $D=5\times10^{24}$\,N\,m is the lithosphere's flexural rigidity, and $\sigma_{zz}$ corresponds to the ice load \citet{bueler-etal-2007}. In the left-hand part of this equation, the first term accounts for mantle relaxation, the second for point-wise isostasy, and the third for elastic flexure. Please refer to \citet{bueler-etal-2007} for a definition of the pseudo-differential operator $|\nabla|$. Due to high mantle viscosity $\nu$, there exist a time lag between the ice sheet growth and the isostatic bedrock response.

@@

Although we do not believe that bedrock elevation must have been at equilibrium with the ice sheet configuration at the LGM, we recognize that through our simplistic palaeoclimatic approach, isostatic effects are another significant source of error on modelled ice surface elevation, which we omitted to discuss in our manuscript.

Through the till friction angle parameter $\phi$, sliding is controlled by bedrock elevation. This control is meant to reflect the presence of lubricating marine sediment at the ice sheet base. Such sediments are expected below current sea level, and, to a lesser extent, in areas that were submerged during former glacial periods, which is reflected in our simulation set-up by a linear increase of $\phi$ between 0 and 200 meters above sea level, where it reaches its maximum value.

We want to clarify here that $\phi$ is computed only once and for the entire simulation length at the beginning of the run, based on modern bedrock elevation. Because the distribution of marine sediments is not likely to change much during a single phase of ice sheet advance and retreat, sliding is not affected by bedrock elevation changes due to isostatic effects in our runs.

>>>>

where $\nu=1\times10^{21}$\,Pa\,s is mantle viscosity, $\rho_r = 3300$\,kg\,m$^{-3}$ is lithosphere density, $D=5\times10^{24}$\,N\,m is the lithosphere's flexural rigidity, and $\sigma_{zz}$ corresponds to the ice load \citet{bueler-etal-2007}. In the left-hand part of this equation, the first term accounts for mantle relaxation, the second for point-wise isostasy, and the third for elastic flexure. Please refer to \citet{bueler-etal-2007} for a definition of the pseudo-differential operator $|\nabla|$. Due to high mantle viscosity $\nu$, there exists a time lag between the ice sheet growth and the isostatic bedrock response.

@@

Although we do not believe that bedrock elevation must have been in equilibrium with the ice sheet configuration at the LGM, we recognize that through our simplistic palaeoclimatic approach, isostatic effects are another significant source of error on modelled ice surface elevation, which we omitted to discuss in our manuscript.
